\addcontentsline{toc}{chapter}{AI Usage}

This section must document the use of Artificial Intelligence (AI) tools
in the project. If AI tools were used at any stage — writing, design,
coding, research, brainstorming, or data analysis — their use must be
clearly described and justified. Describe:

\begin{itemize}
    \item Which AI tools were used (e.g.\ ChatGPT, Copilot).
    \item How the tools were used (e.g.\ brainstorming, improving
          language, summarising information).
    \item Which parts of the project were influenced by AI, if any.
    \item How the AI-generated content was reviewed or modified by
          the student.
\end{itemize}

\noindent
Students remain fully responsible for the final content of the report,
including the correctness of information, code, and analysis. AI tools
must only be used as support and not as a replacement for independent
academic work. \textbf{You must not use AI for:}

\begin{itemize}
    \item Code generation.
    \item Software design decisions.
    \item Test use cases and related artefacts.
    \item Product Backlog and requirements engineering artefacts.
    \item Writing or enhancing report text (feedback and proofreading
          are permitted).
    \item UI generation.
\end{itemize}

\noindent
The use of AI must follow the official guidelines provided by the
University of Southern Denmark. Ensure your use of AI complies with the
rules described at:\\
\url{https://mitsdu.dk/en/mit_studie/kandidat/negot_kandidat/vejledning-og-support/aipaasdu}
