

% --------------------------------
\section{Frontend Implementation}

The frontend of the system was implemented in the UI layer using the Avalonia UI frameworks together with the MVVM (Model--View--ViewModel) architecture.  
The purpose of the frontend was to provide operators and users with an intuitive interface for selecting optimization scenarios, and visualizing optimization results in a structured and user-friendly way.
The implementation focused on separating the UI layer from the Domain layer and Data layer in order to improve scalability, and code readability. 

\subsection{Frontend Architecture}

The frontend was structured into multiple ViewModels and Views:


\begin{itemize}
    \item MainViewModel.cs (MainWindow.axaml)
    \item AssetViewModel.cs (AssetView.axaml)
    \item DashboardViewModel.cs (DashboardView.axaml)
    \item ResultViewModel.cs (ResultView.axaml)
    \item UnitConfigViewModel.cs
    \item ChartBuilder.cs
\end{itemize}
The MainViewModel functioned as the central navigation controller of the application, while the remaining ViewModels handled the logic for their respective pages.  


\subsection{Main Navigation and View Handling}

The MainViewModel served as the central controller for navigation between different pages of the application. 
The ViewModel maintained a CurrentViewModel property, which determined which page was displayed in the main content area. When the user selected a different tab in the navigation menu, the Navigate() method updated the active ViewModel accordingly.
The tabs were stored using the currently selected tab through the SelectedTab property. 


\subsection{Asset Management Interface}


The Asset Management page was implemented to allow users to:

\begin{itemize}
    \item See details of production units
    \item Change between scenaios 
    \item Put units into maintenance
\end{itemize}
The page logic was implemented inside the AssetViewModel.  
Each production unit was represented using a separate UnitConfigViewModel, which stored unit-specific UI state such as maintenance selection, duration, and scheduling information.

\noindent \newline Two independent configuration collections were implemented:

\begin{itemize}
    \item \_scenario1Configs
    \item \_scenario2Configs
\end{itemize}
This design ensured that changes made in one scenario did not affect the other scenario.  


\noindent \newline A maintenance validation was implemented through the RefreshMaintenanceWarning() method.
The system calculated whether selected maintenance schedules could cause insufficient production capacity relative to heat demand.

If maintenance configurations risked creating shortages, warning messages were dynamically generated and displayed to the user.  
This prevented unrealistic operational configurations before optimization was executed.

\subsection{Data Interface}

The Data page provided users with a high-level overview of system information and operational data. The purpose of this page was to give quick access to key system metrics before running the optimization.



\subsection{Result Visualization Interface}

The Result page is used to visualize the optimization output generated by the backend.
The page logic was implemented in the ResultViewModel, while graph generation was separated into the dedicated ChartBuilder class.

\noindent \newline The interface included:

\begin{itemize}
    \item Scenario selection buttons
    \item Winter/Summer selection buttons
    \item Graphs for the optimization
    \item Graph legends
    \item Time-based axes and labels
\end{itemize}
Implemented graphs:

\begin{itemize}
    \item Heat production
    \item Heat demand
    \item CO$_2$ emissions
    \item Net production cost
    \item Electricity price (For Scenario 2)
\end{itemize} 
The heat production graph used stacked area charts to clearly distinguish production contributions from different units.  
To improve readability, individual units were assigned fixed colors and all graphs shared axis configuration methods.  
The graphs dynamically updated whenever the user changed scenario or season selection.





\subsection{Communication Between Frontend and Backend}

The frontend communicated with backend through the ViewModels. The UI itself did not directly perform optimization calculations instead it sent information user selected to the backend.

\noindent \newline When the user navigated to the Result page, the MainViewModel collected:

\begin{itemize}
    \item Selected scenario
    \item Selected season
    \item Selected production units
    \item Maintenance configurations
\end{itemize}

\noindent \newline These values were then passed into the ResultViewModel constructor:

\begin{verbatim}
CurrentViewModel = new ResultViewModel(
    _optimizer,
    AssetPage,
    scenario,
    isSummer,
    selectedUnits
);
\end{verbatim}

\noindent \newline The ResultViewModel then called the backend optimization methods:

\begin{itemize}
    \item RunScenario1()
    \item RunScenario2()
\end{itemize}
The optimization output was transformed into chart-ready data structures through the ChartBuilder helper methods.
This implementation ensured loose coupling between UI components and the backend.


% --------------------------------
\section{Backend Implementation}

The backend was implemented into the Domain layer of the 3-layer architecture. The backend was responsible for handling optimization logic, maintenance scheduling, production unit management, and processing operational data used by the frontend UI layer.

\noindent \newline The backend structure consisted of the following classes:

\begin{itemize}
    \item Optimizer.cs
    \item ProductionUnit.cs
    \item MaintenanceCalculation.cs
    \item MaintenancePeriod.cs
    \item OptimizationResult.cs
    \item ResultFormatter.cs
\end{itemize}

\subsection{Backend Architecture and Responsibilities}

The backend followed a layered structure where each class had a dedicated responsibility.

\subsubsection{Optimizer Class}

The Optimizer class was responsible for running optimization scenarios and processing operational data.

\noindent \newline Its primary responsibilities included:

\begin{itemize}
    \item Loading seasonal operational data
    \item Managing production units
    \item Calculating production costs
    \item Running optimization scenarios
    \item Generating optimization results
\end{itemize}
The optimizer received operational data from the Data layer through dependency injection:

\begin{verbatim}
public Optimizer(ISourceDataManager sourceDataManager,
                 IAssetManager assetManager)
\end{verbatim}
\hbox{This implementation reduced coupling between system components and improved maintainability.}
The optimizer maintained separate dictionaries for seasonal datasets:

\begin{itemize}
    \item Summer
    \item Winter
\end{itemize}
Each dataset contained hourly operational values:

\begin{itemize}
    \item Heat demand
    \item Electricity prices
\end{itemize}

\subsubsection{ProductionUnit Class}

The ProductionUnit class represented individual heat production units within the system.

\noindent \newline Each production unit contained:

\begin{itemize}
    \item Maximum heat production capacity
    \item Electricity generation
    \item Production costs
    \item Energy consumption
    \item CO$_2$ emissions
\end{itemize}
Availability checks were implemented using the IsAvailable() method:

\begin{verbatim}
public bool IsAvailable(DateTime time)
\end{verbatim}
This method ensured that units under maintenance were excluded form the optimization.

\subsubsection{MaintenanceCalculation Class}

\hbox{The MaintenanceCalculation class handled maintenance scheduling for production units.}

The class created maintenance periods dynamically based on user-defined parameters from the frontend interface.

\noindent \newline The scheduling logic:

\begin{itemize}
    \item Located the selected production unit
    \item Created a maintenance period
    \item Assigned the maintenance interval to the unit
\end{itemize}


\subsubsection{ResultFormatter Class}

The ResultFormatter class was responsible for generating readable text-based optimization reports.

\noindent \newline The implementation formatted optimization results into structured summaries containing:

\begin{itemize}
    \item Selected optimization scenario
    \item Seasonal configuration
    \item Maintenance schedules
    \item Hourly production schedules
\end{itemize}



\subsection{Business Logic Implementation}

The backend implemented two primary optimization scenarios.

\subsubsection{Scenario 1 -- Production Cost Optimization}

Scenario 1 optimized heat production by production cost.

\noindent \newline Optimization process:

\begin{enumerate}
    \item Retrieved current heat demand
    \item Filtered available production units
    \item Sorted units by production cost
    \item Allocated heat production sequentially
    \item Continued until demand was satisfied
\end{enumerate}

\subsubsection{Scenario 2 -- Net Production Cost Optimization}

Scenario 2 extended the optimization logic by including electricity prices into the calculation.

\noindent \newline The net production costs are calculated using:

\begin{equation}
\mathrm{Net\ Production\ Cost} =
\mathrm{Production\ Cost} -
\left(
\frac{\mathrm{Electricity\ MW}}
{\mathrm{Maximum\ Heat\ MW}}
\times
\mathrm{Electricity\ Price}
\right)
\end{equation}
This allowed the system to prioritize units capable of generating value through electricity production.
The implementation dynamically recalculated unit ranking for every hour based on current electricity prices.

\subsection{Data Processing and Validation}

The backend performed several validation and processing operations to ensure stable system behavior.

\noindent \newline Examples included:

\begin{itemize}
    \item Validation of production unit capacity
    \item Verification of available source data
    \item Maintenance availability checking
    \item Demand allocation calculations
\end{itemize}
For example, the optimizer validated heat production capacity before performing calculations:

\begin{verbatim}
if (unit.MaxHeatMW <= 0)
{
    throw new ArgumentException(...);
}
\end{verbatim}
Similarly, the system verified that operational data existed for the requested time before running the optimization.

\subsection{Communication with the Data Layer}

The backend communicated with the Data layer through interfaces:

\begin{itemize}
    \item ISourceDataManager
    \item IAssetManager
\end{itemize}
These interfaces provided access to:

\begin{itemize}
    \item Seasonal operational datasets
    \item Production unit configurations
    \item Asset information
\end{itemize}
The actual loading of data was separated into the Data layer. 
This separation improved modularity and allowed backend optimization logic to remain independent from data layer.


% --------------------------------
\section{Data Management}

The data management was implemented into the Data layer of the 3-layer architecture. It was responsible for managing operational datasets, production unit configurations, and structured communication between the backend and external data.


\noindent \newline The Data layer was divided into three primary components:

\begin{itemize}
    \item Asset Management
    \item Source Data Management
    \item Result Data Management
\end{itemize}


\subsection{Data Layer Structure}

The Data layer consisted of the following main classes and interfaces:

\begin{itemize}
    \item AssetManager.cs
    \item IAssetManager.cs
    \item SourceDataManager.cs
    \item ISourceDataManager.cs
    \item ResultDataManager.cs
    \item IResultDataManager.cs
    \item DataPoint.cs
\end{itemize}
The layer also included external CSV files used as operational datasets:

\begin{itemize}
    \item SummerSourceDataSheet.csv
    \item WinterSourceDataSheet.csv
\end{itemize}


\subsection{Asset Manager}

The AssetManager class handled storage and retrieval of production unit information used by the optimizer.
It also implemented the IAssetManager interface and maintained a collection of ProductionUnit objects.

\noindent \newline Each production unit contained operational parameters such as:

\begin{itemize}
    \item Unit name
    \item Maximum heat production
    \item Electricity production
    \item Energy consumption
    \item Production cost
    \item CO$_2$ emissions
    \item Image path for frontend visualization
\end{itemize}
An example production unit definition was implemented as:

\begin{verbatim}
new ProductionUnit
{
    Name = "GB1",
    MaxHeatMW = 3.0,
    ProductionCost = 510,
    CO2Emissions = 132
}
\end{verbatim}
The AssetManager provided access to all production uinits through:

\begin{verbatim}
public List<ProductionUnit> GetProductionUnits()
\end{verbatim}


\subsection{Source Data Manager}

The Source Data Manager handled loading and processing of operational datasets from CSV files.

\noindent \newline The SourceDataManager class implemented the ISourceDataManager interface and loaded two datasets:

\begin{itemize}
    \item Summer operational data
    \item Winter operational data
\end{itemize}
The datasets were stored internally as dictionaries:

\begin{verbatim}
Dictionary<DateTime, DataPoint>
\end{verbatim}
Each timestamp mapped to a DataPoint object containing:

\begin{itemize}
    \item Heat demand
    \item Electricity price
\end{itemize}
The DataPoint model was implemented as:

\begin{verbatim}
public class DataPoint
{
    public double HeatDemand { get; set; }
    public double ElectricityPrice { get; set; }
}
\end{verbatim}
This structure enabled efficient lookup of operational values during optimization.

\subsection{CSV Data Loading}

Operational datasets were loaded from CSV files during application startup.

\noindent \newline The backend resolved file locations dynamically using:

\begin{verbatim}
ResolveSourceDataPath()
\end{verbatim}

\hbox{This implementation improved portability by supporting multiple deployment paths.}

\noindent \newline The CSV loading process included:

\begin{enumerate}
    \item Opening the selected file
    \item Skipping the header row
    \item Reading each line sequentially
    \item Parsing timestamps and numeric values
    \item Creating DataPoint objects
    \item Storing the results in dictionaries
\end{enumerate}
The parsing logic converted the CSV values into structured data objects:

\begin{verbatim}
DateTime time = DateTime.Parse(parts[0]);
double heat = double.Parse(parts[1]);
double price = double.Parse(parts[2]);
\end{verbatim}
Which were later used in the optimization.

\subsection{Result Data Management}

The Result Data Management component handled storage and formatting of optimization results. 

\noindent \newline The generated CSV data included:

\begin{itemize}
    \item Scenario number
    \item Seasonal configuration
    \item Timestamp
    \item Production unit name
    \item Heat production
    \item CO$_2$ emissions
\end{itemize}
The CSV structure was generated using:

\begin{verbatim}
csv.AppendLine(
"Scenario,Season,Hour,Unit,HeatMW,Co2KgPerMWh");
\end{verbatim}
The implementation stored the generated output using:

\begin{verbatim}
public string LastSavedCsv { get; private set; }
\end{verbatim}
This allowed optimization results to be displayed in the UI.

\subsection{Data Validation and Consistency}

Several validation checks were implemented to ensure stable and consistent data handling.

\noindent \newline Examples included:

\begin{itemize}
    \item Verification that CSV files existed before loading
    \item Exception handling for invalid numeric values
    \item Consistent CSV result formatting  
\end{itemize}
If a dataset file could not be found, the system generated an exception:

\begin{verbatim}
throw new FileNotFoundException(...)
\end{verbatim}
This approach prevented the program from completly crashing if the file is not found.

\subsection{Integration with the Backend}

The Data layer supplied operational datasets and production unit information to the backend.

\noindent \newline This allowed the backend to access:

\begin{itemize}
    \item Seasonal operational datasets
    \item Production unit configurations
    \item Structured optimization results
\end{itemize}




